\textls[-0]{\thetitle}

### 2023-04-01 introduction.tex
hyphen

There is a growing trend among programming languages researchers
to use proof assistants to mechanize meta\-theories.
%
However, the programmer runs into an old problem
in the new setting of proof engineering:
the \emph{expression problem} (EP) \cite{wadler-ep}.

@@

The EP is well studied in the setting of ordinary functional
programming and object-oriented (OO) programming.
Modern languages, such as Scala \cite{scala-oopsla05}, have a good
supply of linguistic features that offer expressive power to solve the
EP. %\rD{There is no shortage of solutions to the EP in various
%languages, but there are no citations to those works.}

@@

%\noindent
%As an example, consider \cref{fig:STLC-example}, the simply typed
%lambda calculus (STLC) mechanized in Coq.
%A programming challenge there is to define distinct extensions of this
%STLC formalization to support distinct features,
%and then selectively compose these extensions to form new STLC variants.
As an example, the diagram to the right depicts an extensible
mechanization of the simply typed λ-calculus (STLC), using family
polymorphism.
An extension of STLC with products and another with fixpoints
can both inherit from the base STLC family:
they reuse mechanized metatheories,
from abstract syntax all the way to the type-safety theorem,
only adding new constructors to inductive types
and adding new cases to recursive functions and induction proofs
\emph{as needed} by an extension.

@@

\item We capture the essence of the new language mechanism formally by extending
Martin–Löf type theory with facilities to express family polymorphism (\cref{sec:metatheory2}).
We prove foundational meta\-theoretical results including consistency
and canonicity.
%\EDJ{I will suggest use ``foundational'' instead. Sterling at some places say canonicity is the most basic meta-theoretical property.}
%We also formalize the translation implemented by the Coq plugin.

>>>>

There is a growing trend among programming languages researchers
to use proof assistants to mechanize meta\-theories.
%
However, the programmer runs into an old problem
in the new setting of proof engineering:
the \ac{EP} \cite{wadler-ep}.

@@

The EP is well studied in the setting of ordinary functional
programming and \ac{OO} programming.
Modern languages, such as Scala \cite{scala-oopsla05}, have a good
supply of linguistic features that offer expressive power to solve the
EP. %\rD{There is no shortage of solutions to the EP in various
%languages, but there are no citations to those works.}

@@

%\noindent
%As an example, consider \cref{fig:STLC-example}, the simply typed
%lambda calculus (STLC) mechanized in Coq.
%A programming challenge there is to define distinct extensions of this
%STLC formalization to support distinct features,
%and then selectively compose these extensions to form new STLC variants.
As an example, the diagram to the right depicts an extensible
mechanization of the \ac{STLC}, using family
polymorphism.
An extension of STLC with products and another with fixpoints
can both inherit from the base STLC family:
they reuse mechanized metatheories,
from abstract syntax all the way to the type-safety theorem,
only adding new constructors to inductive types
and adding new cases to recursive functions and induction proofs
\emph{as needed} by an extension.

@@

\item We capture the essence of the new language mechanism formally by extending
Martin-Löf type theory with facilities to express family polymorphism (\cref{sec:metatheory2}).
We prove foundational meta\-theoretical results including consistency
and canonicity.
%\EDJ{I will suggest use ``foundational'' instead. Sterling at some places say canonicity is the most basic meta-theoretical property.}
%We also formalize the translation implemented by the Coq plugin.
